---
title: "Relation between population statistics and sample statistics"
author: "Karl Andersson"
date: "Wednesday, March 11, 2015"
output: html_document
---

##Overveiw

This report outlines some facts about the the relationship a population's statistics and a the statistics of a sample of the populatio. A simulation is made to empiracally prove these relationships.

##Simulations

Simulations were carried out by generating the mean of 40 observations drawn from a exponantially distributed population with lamda = 0.2. This operation is repeated 1000 times to get a vector with 1000 sample means.

```{r}
lambda = 0.2
mns= NULL
for (i in 1 : 1000) {
  mns= c(mns, mean(rexp(40,lambda)))
}
```

##Sample Mean versus Theoretical Mean
We can se that the the mean of the sample means are very close to the theoretical mean of the pupulation.Remember that lambda was set to 0.2.

```{r}
cat(sprintf('Theoretical Mean of Distribution= 1/ lamda = %f', 1/lambda))
cat(sprintf('Sample mean(1000 samples) = %f', mean(mns)))
```


##Sample Variance versus Theoretical Variance

The variation of the sample mean is much smaller than the variance of the total population 

```{r}
cat(sprintf('Sample variance = 1/ lamda = %f', var(mns)))
cat(sprintf('Teoretical variance of distribution = (1/lamda)^2 = %f' ,((lambda)^-2)))

cat(sprintf('Population variance / Sample variance = %f', (lambda^-2)/var(mns)))
cat(sprintf('Square root of 1000 = %f',sqrt(40)))
```

##Sample Mean aprox. normal distributed

The Means are normally distributed, which can be seen as the the frequenzy function is bell shaped and symmetric around the mean.
```{r}

hist(mns)
```

THe original enponential distribution is not normally distributed
```{r}
hist(rexp(40,0.2), breaks = 50)
```
This is seen even better with a histogram of more sample points

```{r}
hist(rexp(40000,0.2), breaks = 50)
```