---
title: "Prediction Movement Data"
author: "Karl Andersson"
date: "Friday, April 24, 2015"
output: html_document
---

#Preparation and Method
###Load Libraries and data
```{r message=FALSE, warning=FALSE}
require(dplyr)
require(caret)
require(randomForest)
training = read.csv("pml-training.csv", header = T)
test = read.csv("pml-testing.csv", header = T)
```

###Data Cleaning

Remove Outcome variable before transformation

```{r}
classe <- training$classe
```

remove columns with all NA-values and X-column
```{r}
training = training[,colSums(is.na(training)) ==0]
training = select(training,-X)
```

Select the same columns from test as from training
```{r}
test <- test[,which(names(test) %in% names(training))]
```

Convert columns to numeric
```{r}
training <- training[, sapply(training, is.numeric)]
test <- test[,sapply(test, is.numeric)]
```

add the outcome variable which was removed before cleaning
```{r}
training$classe <- classe
```

###Splitting Data
Set seed and aprtition data into training and validation sets
```{r}
set.seed(12312)
trainIndex <- createDataPartition(training$classe, p=0.70, list=F)
trainData = training[trainIndex,]
validationData = training[-trainIndex,]
```

###Train Model
Train a randomForest predictor with classe as outcome and all other columns as predictors. 
```{r cache = TRUE}
train_control <- trainControl(method="boot", number=3,allowParallel=T)
model <- train(as.factor(classe) ~., data = trainData, method = "rf",trControl=train_control)
```
###Make Predictions

Make Predictions on Training and Validation set using the trained model
```{r}
predictTrain <- predict(model,trainData)
predictValidation <- predict(model,validationData)
```

#Analysis and results
###Confusion Matrix
A confusionmatrix on the validation predictions and validation set shows what what prediction the model makes correct and which not. The accurancy is very high 0.9997.
```{r}
confusionMatrix(predictValidation,validationData$classe)
```

Our out-of test was 0.00034 = 0.034%
```{r}
valInd <- validationData$classe == predictValidation
ErrorVal = length(valInd[valInd == FALSE])/ length(valInd)
ErrorVal
```
#Prediction of test Dataset
```{r}
predictTest = predict(model,newdata=test)
predictTest
```